---
title: "BSG HW3"
author: "Sten-Oliver Salumaa, Denis Kovalenko"
date: "13 December 2017"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Task1
The file ABO-CHB.rda contains genotype information of individuals of a Chinese population of unrelated
individuals. The genotype information concerns SNPs the ABO bloodgroup region, located
on chromosome number 9. The file contains genotype information (Z, individuals in columns, SNPs
in rows), the physical position of each SNP (pos) and the alleles for each SNP (alleles). Load
this data into the R environment

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the data
library(genetics)
library(HardyWeinberg)
filename = "ABO-CHB.rda"
load(filename)
```


## Task2 (1p)
How many individuals and how many SNPs are there in the database?

```{r echo=FALSE}
print("Individual count:")
ncol(Z)
print("SNP count:")
nrow(Z)
```
What percentage of the data is missing?

```{r echo=FALSE}
total <- ncol(Z)*nrow(Z)
nas <- length(Z[is.na(Z)])
nas/total*100
```

## Task3 (1p)
Depict all SNPs simultaeneously in a ternary plot, and comment on your result. Do you
believe Hardy-Weinberg equilibrium is tenable for the markers in this database?

```{r echo=FALSE}
# transpose the data to make it more logical
Zt <- t(Z)

# create a map function for groupgenotypes
map <- list("AA"=c("A/A"),
            "TT"=c("T/T"),
            "CC"=c("C/C"),
            "GG"=c("G/G"),
            "AB"=".else")

geno_loci <- data.frame(AA = numeric(0), AB = numeric(0), BB = numeric(0))
for(i in 1:ncol(Zt)){
  # genotype counts table as AA, AB, and BB
  #print(expectedGenotypes(genotype(Zt[,i], sep=""), ploidy=2))
  genocounts <- table(groupGenotype(x=genotype(Zt[,i], sep=""), map=map, factor=FALSE))
  #print(genocounts)
  #print(names(genocounts))
  #passing genotype counts to table
  if(length(genocounts)==3){
    geno_loci[i,"AB"] <- genocounts["AB"]  
    genocounts <- genocounts[names(genocounts) != "AB"]
    geno_loci[i, "AA"] <- genocounts[1]
    geno_loci[i, "BB"] <- genocounts[2]
  }else if(length(genocounts)==2){
    geno_loci[i,"AB"] <- genocounts["AB"]  
    genocounts <- genocounts[names(genocounts) != "AB"]
    geno_loci[i, "AA"] <- genocounts[1]
  }else if(length(genocounts)==1){
    geno_loci[i,"AB"] <- genocounts["AB"]  
    genocounts <- genocounts[names(genocounts) != "AB"]
    geno_loci[i, "AA"] <- genocounts[1]
  }

}

# replace n.a.-s
geno_loci[is.na(geno_loci)] <- 0
# plot the thing
HWTernaryPlot(geno_loci)
```

Yes, it seems that HW equilibrium is tenable for this database since almost all of the data points reside in the equilibrium boundaries.




```{r echo=FALSE}



```

```{r echo=FALSE}



```


```{r echo=FALSE}



```


```{r echo=FALSE}



```


```{r echo=FALSE}



```


```{r echo=FALSE}



```

```{r echo=FALSE}



```